\section{Project Requirements Checklist}

\textbf{Computer Components}

\begin{itemize}
    \item 1 Raspberry Pi per each of the following: Smoke Detector device, Smoke Alarm, Wearable Accessory
    \item 1 Raspberry Pi will be used to transfer info to and from all the other RPis to the Web GUI (running in headless mode). There are 4 Raspberry Pi’s used in the project overall.
\end{itemize}

\textbf{Hardware Components}

\begin{itemize}
    \item MQ2 Smoke Sensor (x1), LCD Display (x2), AudioHat Module (x1), SenseHat Module (x1)
    \item Actuator: LCD Display, AudioHat Module
    \item Smoke sensor waits till it senses smoke, once that requirement is met it triggers the alarm, wearable accessory, and
          sends a notification to the Web GUI
\end{itemize}

\textbf{Software Components}

\begin{itemize}
    \item Database for employee contact information.
    \item The computer hosting the database changes depending on user configurations (Level of smoke to detect, Web GUI themes,
          alarm sounds).
    \item Database includes a table for user info (employee names, phone numbers, emergency numbers) and another table for sensor
          data (temperature data, smoke detection).
    \item Periodic timing loop is given by the smoke sensor monitoring the environment for smoke. The smoker sensor will log this
          data and send it to the web GUI every minute.
    \item The data the smoke detector logs will be displayed on the user website in the form of varying graphs and widgets.
    \item Smart watch will have a buzzer, Raspberry Pi with the alarm will have notification software. Raspberry Pi will send
          notifications to the Smart watch and to the web GUI.
    \item Bidirectional GUI included (user can reset the smoke detector from the web GUI, and changes made to the physical
          detector will show in Web GUI settings)
\end{itemize}
