\section{Work Plan}

The following section describes the approach taken by the FANS development team to collaborate effectively and complete
the FANS system before its deadline.

\subsection{The Project Team}

Each member of the project team is a third year computer systems engineering student.

\textbf{Grant Achuzia} \\
Grant's strengths lie in effective project management and organization, and he is passionate about delivering successful
projects in academic settings. Grant also has significant experience designing web interfaces that are accessible and
aesthetically pleasing.

\textbf{Matteo Golin} \\
Matteo has background in developing embedded system on Raspberry Pi's from his work at Carleton University's rocketry
design team, where he is leading the development of a real-time telemetry system using Blackberry's QNX RTOS.

\textbf{Saja Fawagreh} \\
Saja has developed skills in user interface design from her time on co-op, and excels and making visually appealing
interfaces because of her sharp attention to detail.

\textbf{Javeria Sohail} \\
Javeria has expertise in system architecture design, and has a strong knowledge of object-oriented design patterns. She
is passionate about optimization, efficient design and modularity.

\subsubsection{Roles and Tasks}

% TODO: modify responsibilities
\begin{table}[H]
    \centering
    \begin{tabular}{| c | p{3cm} | p{4cm} | p{5cm} |}
        \hline
        \textbf{Team Member} & \textbf{Major Role/Task (Primary)} & \textbf{Secondary Responsibility (Testing)}        &
        \textbf{Reason for assignment}                                                                                                                                                                              \\
        \hline
        Grant Achuzia        & Project coordination               & Organize the entire system (software and hardware) & Strong project
        management and organizational skills, ensuring academic milestones are met on time.                                                                                                                         \\
        \hline
        Matteo Golin         & Software development               & Testing and debugging                              & Skilled in programming and problem-solving, enthusiastic about tackling coding challenges. \\
        \hline
        Saja Fawagreh        & UI design                          & Front-end testing and usability                    & Proficient in front-end
        development, dedicated to creating user-friendly interfaces.                                                                                                                                                \\
        \hline
        Javeria Sohail       & Project architecture               & hardware integration                               &
        Familiar with system architecture and design for hardware components, ensuring seamless hardware functionality
        within FANS.                                                                                                                                                                                                \\
        \hline
    \end{tabular}
    \caption{Assignment of team members to major project tasks}
\end{table}

\subsubsection{Teamwork Strategy}

\textbf{Regular Team Meetings} \\
We will schedule regular team meetings to discuss project progress, share updates, and address any challenges or
questions. Our team will establish MS Teams as our main communication channel to ensure efficient communication within
the team and with the TA. Additionally, the team meetings will be documented to review what we talked about in case
someone missed the meeting.

\textbf{Version Control} \\
We will use GitHub to manage codebase changes. This includes practices such as creating branches for features or bug
fixes, committing regularly, and using pull requests for code reviews. Informative commit messages will be prioritized
to avoid misunderstandings.

\textbf{Code Reviews} \\
Before merging code changes, thorough code reviews will be conducted to ensure code quality and maintainability.

\textbf{Testing} \\
We will follow a testing strategy that includes unit testing, integration testing, and system testing to ensure the
reliability and robustness of the FANS.

\textbf{Feedback} \\
Team members will provide constructive feedback to each other for continuous improvement and learning (also making use
of FBF). We also want to take constant feedback from our TA at every milestone so that our big picture comes together
seamlessly.

\subsubsection{What We Will Need to Learn}

\textbf{Hardware integration} \\
Since the FANS includes both hardware and software components, we need to learn how to integrate them seamlessly. We
utilize online resources and forums where hardware integration techniques are discussed.

\textbf{Real-time communication} \\
Implementing real-time communication features such as SMS and email notifications requires
knowledge of APIs and protocols for sending messages. We will look at the documentation and instructions provided by
communication service providers (e.g. SMS) and learn how to integrate these services into our system.

\textbf{Developing a web interface} \\
Creating a user-friendly web interface for the FANS requires knowledge of front-end development, including HTML, CSS
and JavaScript. We will use online tutorials, courses and documentation of web development frameworks such as React or
Angular to improve our web development skills.

\textbf{Test strategies} \\
Developing effective testing strategies, including unit testing, integration testing, and system
testing, requires an understanding of various testing frameworks and methodologies. We will explore tutorials,
documentation, and best practices for testing in software development.

\subsection{Project Milestones}

% TODO: This seems very unrealistic %
% TODO: Shouldn't sensor integration go before smoke detection
\begin{table}[H]
    \centering
    \begin{tabular}{| c | c | p{9cm} |}
        \hline
        \textbf{Milestone Name} & \textbf{Date} & \textbf{Description}                                                                            \\
        \hline
        UI                      & 2024-02-13    & Develop a user-friendly web interface for monitoring system status, configuring thresholds, and
        interacting with the system.                                                                                                              \\
        \hline Database         & 2024-02-17    & Set up a database to store employee contact information
        and system data, ensuring efficient management and retrieval.                                                                             \\
        \hline Smoke detection  & 2024-02-20    & Develop and test
        the smoke detection algorithm on the Raspberry Pi to ensure accurate detection of fire and smoke.                                         \\
        \hline
        Sensor integration      & 2024-03-02    & Integrate sensors with the Raspberry Pi for real-time data collection, enabling the
        system to monitor environmental conditions.                                                                                               \\
        \hline
        Notifications           & 2024-02-12    & Implement a notification system to trigger alarms and alert users via SMS and email
        in case of fire emergencies.                                                                                                              \\
        \hline
        Testing                 & 2024-02-20    & Conduct comprehensive testing to ensure the functionality and integration of both
        hardware and software components.                                                                                                         \\
        \hline
        Finalization            & 2024-02-27    & Finalize system configuration, including thresholds and notification settings, and
        conduct comprehensive testing.                                                                                                            \\
        \hline
    \end{tabular}
    \caption{Major milestones for FANS development.}
\end{table}

\subsection{Schedule of Activities}

Under construction!

% TODO: Requires finished figure from Grant
